Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-08-31 Manuscript.Rmd
Fixing figures mainly
manuscrip edits

To quantify the agreement between partitions, we computed both the adjusted rand index and variation of information distance. The adjusted rand index denotes the proportion of vertices that coincide in affiliation across partitions, and compares this score to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).

However, this time I also used the rand index to quantify the community correspondence between subjects. Specifically, I divided each subject’s spectrally- partitioned brain into PCC and mPFC topographical zones (posterior and frontal brain, respectively), and computed the index of each region between subjects. The high cross-subject heterogeneity of cortical folding of mPFC should produce relatively low rand indices, while the more homogeneous PCC should display the opposite pattern. This analysis was meant to tackle the subject-specific limitations discussed above. The final patterns were inspected visually.

@@

This first step just gives an idea of the level of agreement within and between subjects. As a first measure, we use the adjusted rand index (RI) to estimate the level of agreement between partitions. RI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes (e.g. clustered together in partition 1, but separately in 2). It is worth noting that the adjusted rand index has well known limitations (Fortunato & Hric, 2016). To compensate for this, we also performed all the upcoming analyses with the variation of information index (VI). This distance measure denotes how much information is gained by adding the second partition (thus, lower values are better), and is known for its robustness. Both RI and VI analyses agreed in the results they yielded, but we only report RI due to its easier interpretability. This first plot shows the distribution of RI values between modularity and spectral partitioning for each subject.

@@

The total number of subjects is `r nSubj`. The distribution suggests that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)` sd = `r round(sd(RI_all_within$Index), digits = 2)`). The importance of this results lies in that while spectral partitioning forces a bipartition of the network, modularity does not. This adds evidence to the division of these areas into DMN and non-DMN during rest. Due to this high agreement, we focus on results using spectral partitioning, as it paints a clearer picture of the division of interest.

>>>>

In order to evaluate the stability and topographical heterogeneity of the partitions within and across subjects, we chose the adjusted rand index (RI; Rand, 2012) as the reference metric of agreement between two clusterings. The adjusted rand index estimates the rate at which every pair of nodes in partition A were equally assigned to the same/different parcels on partition B, over all possible changes (e.g. clustered together on A, but separately on B). The resulting ratio is compared to a baseline given by the expectation based on a random vertex assignment for an equal number of clusters across partitions. Because of this, the adjusted rand index is a simple metric that determines the percentage agreement of any two partitions that is agnostic to the labeling scheme. *(SHOULD WE INCLUDE THIS?) On the other hand, variation of information gives a general sense of the amount of information to be gained by the complementary community partitioning (Meila, 2007), measured by the addition of the conditional entropies per combination of clusters across partitions. While the rand index is more easily interpreted, it suffers from the assumption that the baseline should have an equal number of vertices per cluster across partitions, which is not necessarily the case. Therefore, I used variation of information as a more robust similarity metric (Fortunato & Hric, 2016).*

The topographies were compared: 1) Across community detection methods; 2) Overall data across individuals (ADD); 3) Overall data across participants, PCC and mPFC separately; 4) Day 1 vs Day 2 overall; 5) Day 1 vs Day 2 PCC vs mPFC; 6) Across sessions overall; 7) Across sessions PCC vs mPFC; 8) Sliding windowper subject overall; 9) Sliding window PCC vs mPFC. We compared the agreement within subjects against across subjects by computing the ratio of the mean RI within and between subjects. Ratios close to one denote good alignment across subjects, while ratios over one mean that partitions per subject were more stable than what you would expect against every other subject.

@@

This first step just gives an idea of the level of agreement within and between subjects. As a first measure, we use the adjusted rand index (RI) to estimate the level of agreement between spectral partitioning and modularity. RI measures the proportion of pairs of nodes in a network that were either clustered together or separately in both partitions, over all possible affiliation changes (e.g. clustered together in partition 1, but separately in 2). It is worth noting that the adjusted rand index has well known limitations (Fortunato & Hric, 2016). To compensate for this, we also performed all the upcoming analyses with the variation of information index (VI). This distance measure denotes how much information is gained by adding the second partition (thus, lower values are better), and is known for its robustness. Both RI and VI analyses agreed in the results they yielded, but we only report RI due to its easier interpretability. This first plot shows the distribution of RI values between modularity and spectral partitioning for each subject.

@@

The total number of subjects is `r nSubj`. The distribution suggests that both clustering methods were in high agreement across subjects (mean = `r round(mean(RI_all_within$Index), digits = 2)` sd = `r round(sd(RI_all_within$Index), digits = 2)`). The importance of this results lies in that while spectral partitioning forces a bipartition of the network, modularity does not. This adds evidence to the division of these areas into DMN and non-DMN during rest.
